---
title: "Weekly Summary Template"
author: "Brady Miller"
title-block-banner: true
title-block-style: default
toc: true 
format: html
# format: pdf
---

```{r}
library(tidyverse)
library(ISLR2)
library(cowplot)
library(kableExtra)
```

## Tuesday, February 7th

::: {.callout-important}
## TIL

Include a _very brief_ summary of what you learnt in this class here. 

Today, I learnt the following concepts in class:

1. What the interpretation of $\beta_0\ and $\beta_1$ is (regression coefficients)
1. Categorical covariates
1. How to reorder factors
:::

#### *Interpretation of $\beta_0\ and $\beta_1$ (regression coefficients)*

* The regression model is...

$y_i = \beta_0 + \beta_1 x_i + \epsilon_i$

* $\beta_0$ and $\beta_1$ are the regression coefficients with $\beta_0$ being the intercept and $\beta_1$ being the slope

```{r}
library(ggplot2)
attach(mtcars)

x <- mtcars$hp
y <- mtcars$mpg

model <- lm(y ~ x)
summary(model)
```
Based on the summary model...

* The intercept means that a car with 'hp=0' would have 'mpg=30.09'
## Thursday, February 9th


* From this, if we have some co-variate $x_0$, then the expected value for $y(x_0)$ is given by:

$$y(x_0) = \beta_0 + \beta_1 x_0$$

* Using this we can find the expected value for $x_0 + 1$, which is...

$$
\begin{align}
y(x_0 + 1) = \beta_0 + \beta_1 \times (x_0 + 1)\\ \\
&= \beta_0 + \beta_1 x_0 + \beta_1\\ \\
&= y(x_0) + \beta_1\\ \\above

\implies \beta_1 &= y(x_0 + 1) - y(x_0)
\end{align}
$$

#### *Categorical covariates*
In class, we looked at categorical covariates in the iris dataset, and created a boxplot and a summary of the model for the categorical covariates

```{r}
x <- iris$Species
y <- iris$Sepal.Length

# boxplot(Sepal.Length ~ Species, df)
boxplot(y ~ x)
```

```{r}
cat_model <- lm(Sepal.Length ~ Species, iris)
cat_model
```

Even if $x$ is categorical, we can still write down the regression model as follows:
$$
y_i = \beta_0 + \beta_1 x_i
$$
where $x_i \in \{setosa, \ versicolor, \ virginica \}$. This means that we end up with, three different models with each one having a different intercept.

1. $y_i = \beta_0 + \beta_1 (x_i == 'setosa')$
1. $y_i = \beta_0 + \beta_1 (x_i == 'versicolor')$
1. $y_i = \beta_0 + \beta_1 (x_i == 'virginica')$

* The interpretation of the intercept ($\beta_0$) is the expected y value when x belongs to the base category
* The slope ($\beta_1$) with the name 'Species.versicolor' represents the following:

* '(Intercept)' = $y(x = \texttt{setosa})$

* 'Species.versicolor' = $y(x = \texttt{versicolor}) - y(x = \texttt{setosa})$
* 'Species.virginica' = $y(x = \texttt{virginica}) - y(x = \texttt{setosa})$


#### *Reordering factors* 
Lets say that we didn't want 'setosa' to be the baseline level, and instead, we wanted 'virginica' to be the baseline level. How would we do this?

First we reorder/relevel the categorical covariate
```{r}
# before
iris$Species
iris$Species <- relevel(iris$Species, "virginica")

# after
iris$Species
```

Once we do the re-leveling, we can now run the regression model:
```{r}
new_cat_model <- lm(Sepal.Length ~ Species, iris)
new_cat_model
```







## Tuesday, February 9th

::: {.callout-important}
## TIL

Include a _very brief_ summary of what you learnt in this class here. 

Today, I learnt the following concepts in class:

1. Item 1
1. Item 2
1. Item 3
:::

